Constructive Preference Elicitation by Setwise Max-margin Learning

### 2016-04-18 paper/ijcai16.tex
paolos edits

\documentclass{article}
\usepackage{ijcai16}
\usepackage{times}
%\usepackage{latexsym}
%\usepackage{graphicx}
\usepackage{graphbox}
\usepackage{hyperref}
\usepackage{multirow}
\usepackage{placeins}
\usepackage[font=small]{caption}

@@

@@

\author{Stefano Teso \\
University of Trento \\
Trento, Italy \\
teso@disi.unitn.it
\And
Andrea Passerini \\
University of Trento \\
Trento, Italy \\
passerini@disi.unitn.it
\And
Paolo Viappiani \\
Universit\'e Pierre et Marie Curie \\
Paris, France \\
paolo.viappiani@lip6.fr}

@@

\section{Introduction}

\paolo{or Andrea: check whether~\cite{greco2008ordinal} is related and comment
on it, as requested by reviewer 3.}

Preferences \cite{Peintner2008} play an important role in a variety of artificial
intelligence applications and the task of eliciting or learning preferences is a crucial one; 
typically only limited
information about the user's preferences will be available and the
cost (cognitive or computational) of obtaining additional preference information will be high.  
%The goal in
%preference elicitation is that of gaining enough information about the
%user's
%preferences %(typically represented by an utility function $u$, unknown to the system)
%in order to recommend a decision or a course of action.

The automated assessment of preferences has received considerable attention,
starting with pioneering works \cite{White1984,jaquetlsiskos1982} in the OR community.
\paolo{or Andrea, check if it is worth mentioning the review in~\cite{jaquetlsiskos2001},
as highlighted by the fourth reviewer.}
%In the past decade a number of researchers have considered elicitation of preferences from an artificial intelligence perspective%; within the machine learning field the term preference learning has been catching up.
%While preferences are often assessed using some ad-hoc rules \cite{},
More recently, a number of AI researchers have proposed interactive methods that elicit
preferences in an adaptive way
\cite{chajewska2000,boutilier2002,Wang2003,boutilier2006,guo2010real,viappiani2010optimal},
observing that, by asking informative questions, it is often possible to make near-optimal decisions with only partial preference information.

@@

Consistently with previous work~\cite{guo2010real,viappiani2010optimal}, 
we model users with additive utility functions \cite{keeney1976}; 
the user's preferences are represented by a  weight
vector $\vw\in\bbR^m$ and
%We assume linear utility functions \cite{keeney1976}; 
the utility of a configuration $\vx$ is given by
$\langle \vw, \vx \rangle = \sum_{z=1}^m w_z x_z$. 
In the remainder of the paper we
require all weights to be {\em non-negative} and {\em bounded}: the per-attribute
weights $w_z$ must lie in a (constant but otherwise arbitrary) interval
$[w^\bot_z, w^\top_z]$, with $w^\bot_z \ge 0$. 
Both requirements are quite natural\footnote{Utility values are defined on an
interval scale, thus it is always possible to scale the values appropriately (see for instance \cite{Torra2007} and \cite{keeney1976}).}, and enable the translation of our core optimization problem into a
mixed-integer linear problem (as done in Section~\ref{sec:formulation}).

@@

The first goal is achieved by translating all pairwise preferences
$\calD$ into ranking constraints: preferences of the form
$\vy^h_+ \succ \vy^h_-$ become linear inequalities of the form
$\langle \vw^i, \vy^h_+ - \vy^h_- \rangle \geq \mu$, where $\mu$ is the
{\em margin} variable (which we aim at maximizing) and $h$ ranges over
the responses.  Non-separable datasets, which occur in practice due to
occasional inconsistencies in user feedback, are handled by
introducing slack variables (whose sum we aim at minimizing). When
augmented with the slacks, the above inequalities take the form
$\langle \vw^{i}, \vy^{h}_+ - \vy^{h}_- \rangle \ge \mu - \varepsilon^{i}_h$
where $\varepsilon^{i}_h$ is the penalty incurred by weight vector $\vw^{i}$
for violating the margin separation of pair $h$. Indifference preferences, i.e.
$\vy^h_1 \approx \vy^h_2$, are translated as $|\langle \vw^i, \vy^h_1 - \vy^h_2 \rangle| < \varepsilon^i_h$;
the slack increases with the difference between the estimated utility of the
two options.

@@

In order to study the effect of increasing the number of weight
vectors in our formulation, we also ran {\sc SetMargin} varying the
parameter $k$. Figure~\ref{fig:selfcomparison} reports utility loss
results on $r\!=\!4$ and $r\!=\!5$ datasets for the uniform and sparse normal
distributions (the toughest and the simplest, for space limitations).
% \footnote{We picked the most difficult dense distribution
%   (uniform) and the simplest sparse distribution (normal) to provide
%   the broadest possible picture, as space constraints prevent us from
%   including all results. Results for the other distributions are
%   however qualitative similar to the ones reported.}, 
The first and third columns report results in terms of number of
iterations. It can be seen that increasing the number of weight
vectors tends to favour earlier convergence, especially for the more
complex dataset ($r=5$). However, as in each iteration the user is
asked to compare $k$ items, different values of $k$ imply a different
cognitive effort for the user. The second and fourth columns report
results in terms of number of queries, where we count all
$k \choose 2$ pairs of queries when comparing $k$ items. In this case,
$k=2$ seems to be the best option. The cognitive cost for the user
will likely lay in between these two extremes, but formalizing this
concept in an efficient query ordering strategy needs to face the
effect of noise.  A modified sorting algorithm asking only
$O(k\log k)$ queries to the user resulted in a performance worsening,
likely because of a cascading effect of inconsistent feedback (but
could be beneficial with different noise levels).

% \paragraph{Cognitive effort.} Contrarily to most other preference elicitation
% methods, which generate a single candidate option per iteration, {\sc
% SetMargin} constructs a whole {\em set} of options. We thus can not leverage
% standard query selection strategies \stefano{which ones?}.

% The simplest solution is to query the user about all $k \choose 2$ pairs of
% configurations in $\{\vx^i\}$. Since $k$ is a small constant, the resulting
% number of queries does not grow large in practice.  An appealing alternative
% involves using a modified sorting algorithm, such as merge-sort, to order the
% set $\{\vx^i\}$ by asking only $O(n\log n)$ queries to the user. However, this
% approach is very susceptible to noise: a single inconsistent user response may
% affect all successive comparisons, leading to a cascading effect. While more
% robust strategies may be conceived, empirically we found that the simpler
% quadratic scheme tends to work reasonably well.

%\paolo{Maybe a plot of an experiment with increased noise in user response? (for example, $\beta=0.1$ instead of $1$.}

@@

We presented a max-margin approach for efficient preference
elicitation in large configuration spaces.\footnote{Note that max-margin learning has been proposed before \cite{gajos2005} for preference elicitation, 
but with  rudimental methods for query selection.}
Our approach relies on an
extension of max-margin learning to sets, and is effective in the
generation of a diverse set of configurations that can be used to ask
informative preference queries.  The main advantages of this
elicitation method are 1) ability to provide recommendations in large
configuration problems 2) robustness with respect to erroneous
feedback and 3) ability to encourage sparse utility functions.
Experimental comparisons against state-of-the-art Bayesian preference
elicitation strategies confirm these advantages. Future work includes
extending the approach to truly hybrid scenarios (where real valued
attributes do not depend on categorical ones) and studying its
applicability %of the setwise max-margin idea 
to other problems.

%\paolo{any idea about possible future works?}
% FUTURE WORKS

\section*{Acknowledgments}
ST was supported by the Caritro Foundation through project E62I15000530007.

>>>>

\documentclass{article}
\usepackage{ijcai16}
\usepackage{times}
%\usepackage{latexsym}
%\usepackage{graphicx}
\usepackage{graphbox}
%\usepackage{hyperref}
\usepackage{multirow}
\usepackage{placeins}
\usepackage[font=small]{caption}

@@

% TEMPORARY HACK: redefine href as dummy command to avoid Fatal Error!
\newcommand{\url}[1]{#1}
% ---------------------------------------------------------------------------

@@

\author{Stefano Teso \\
University of Trento \\
Trento, Italy \\
teso@disi.unitn.it
\And
Andrea Passerini \\
University of Trento \\
Trento, Italy \\
passerini@disi.unitn.it
\And
Paolo Viappiani \\
Sorbonne Universit\'es\\
UPMC Univ Paris 06\\
CNRS, LIP6 UMR 7606\\
Paris, France \\
paolo.viappiani@lip6.fr}

@@

\section{Introduction}
Preferences \cite{Peintner2008} play an important role in a variety of artificial
intelligence applications and the task of eliciting or learning preferences is a crucial one; 
typically only limited
information about the user's preferences will be available and the
cost (cognitive or computational) of obtaining additional preference information will be high.  
%The goal in
%preference elicitation is that of gaining enough information about the
%user's
%preferences %(typically represented by an utility function $u$, unknown to the system)
%in order to recommend a decision or a course of action.

\paolo{The automated assessment of preferences has received considerable attention,
starting with pioneering works in the OR community, such as \cite{White1984} and especially the UTA methodology \cite{jaquetlsiskos1982} giving rise to a wide variety of extensions \cite{jaquetlsiskos2001,greco2008ordinal}.}

%In the past decade a number of researchers have considered elicitation of preferences from an artificial intelligence perspective%; within the machine learning field the term preference learning has been catching up.
%While preferences are often assessed using some ad-hoc rules \cite{},
Within AI, a number researchers have proposed interactive methods that elicit
preferences in an adaptive way
\cite{chajewska2000,boutilier2002,Wang2003,boutilier2006,guo2010real,viappiani2010optimal},
observing that, by asking informative questions, it is often possible to make near-optimal decisions with only partial preference information.

@@

Consistently with \paolo{the experimental settings} of previous work~\cite{guo2010real,viappiani2010optimal}, 
we model users with additive utility functions \cite{keeney1976}; 
the user's preferences are represented by a  weight
vector $\vw\in\bbR^m$ and
%We assume linear utility functions \cite{keeney1976}; 
the utility of a configuration $\vx$ is given by
$\langle \vw, \vx \rangle = \sum_{z=1}^m w_z x_z$. 
In the remainder of the paper we
require all weights to be {\em non-negative} and {\em bounded}: the per-attribute
weights $w_z$ must lie in a (constant but otherwise arbitrary) interval
$[w^\bot_z, w^\top_z]$, with $w^\bot_z \ge 0$. 
Both requirements are quite natural\footnote{Utility values are defined on an
interval scale, thus it is always possible to scale the values appropriately (see for instance \cite{Torra2007} and \cite{keeney1976}).}, and enable the translation of our core optimization problem into a
mixed-integer linear problem (as done in Section~\ref{sec:formulation}).

@@

The first goal is achieved by translating all pairwise preferences
$\calD$ into ranking constraints: preferences of the form
$\vy^h_+ \succ \vy^h_-$ become linear inequalities of the form
$\langle \vw^i, \vy^h_+ - \vy^h_- \rangle \geq \mu$, where $\mu$ is the
{\em margin} variable (which we aim at maximizing) and $h$ ranges over
the responses.  Non-separable datasets, which occur in practice due to
occasional inconsistencies in user feedback, are handled by
introducing slack variables (whose sum we aim at minimizing)
\paolo{in a way similar to UTA and its extensions \cite{jaquetlsiskos1982,greco2008ordinal} }. When
augmented with the slacks, the above inequalities take the form
$\langle \vw^{i}, \vy^{h}_+ - \vy^{h}_- \rangle \ge \mu - \varepsilon^{i}_h$
where $\varepsilon^{i}_h$ is the penalty incurred by weight vector $\vw^{i}$
for violating the margin separation of pair $h$. Indifference preferences, i.e.
$\vy^h_1 \approx \vy^h_2$, are translated as $|\langle \vw^i, \vy^h_1 - \vy^h_2 \rangle| < \varepsilon^i_h$;
the slack increases with the difference between the estimated utility of the
two options.

@@

In order to study the effect of increasing the number of weight
vectors in our formulation, we also ran {\sc SetMargin} varying the
parameter $k$. Figure~\ref{fig:selfcomparison} reports utility loss
results on $r\!=\!4$ and $r\!=\!5$ datasets for the uniform and sparse normal
distributions (the toughest and the simplest, for space limitations).
% \footnote{We picked the most difficult dense distribution
%   (uniform) and the simplest sparse distribution (normal) to provide
%   the broadest possible picture, as space constraints prevent us from
%   including all results. Results for the other distributions are
%   however qualitative similar to the ones reported.}, 
The first and third columns report results in terms of number of
iterations. It can be seen that increasing the number of weight
vectors tends to favour earlier convergence, especially for the more
complex dataset ($r=5$). However, as in each iteration the user is
asked to compare $k$ items, different values of $k$ imply a different
cognitive effort for the user. The second and fourth columns report
results in terms of number of queries, where we count all
$\binom{k}{2}$
%k \choose 2$ 
pairs of queries when comparing $k$ items. In this case,
$k=2$ seems to be the best option. The cognitive cost for the user
will likely lay in between these two extremes, but formalizing this
concept in an efficient query ordering strategy needs to face the
effect of noise.  A modified sorting algorithm asking only
$O(k\log k)$ queries to the user resulted in a performance worsening,
likely because of a cascading effect of inconsistent feedback (but
could be beneficial with different noise levels).
% \paragraph{Cognitive effort.} Contrarily to most other preference elicitation
% methods, which generate a single candidate option per iteration, {\sc
% SetMargin} constructs a whole {\em set} of options. We thus can not leverage
% standard query selection strategies \stefano{which ones?}. 
% The simplest solution is to query the user about all $k \choose 2$ pairs of
% configurations in $\{\vx^i\}$. Since $k$ is a small constant, the resulting
% number of queries does not grow large in practice.  An appealing alternative
% involves using a modified sorting algorithm, such as merge-sort, to order the
% set $\{\vx^i\}$ by asking only $O(n\log n)$ queries to the user. However, this
% approach is very susceptible to noise: a single inconsistent user response may
% affect all successive comparisons, leading to a cascading effect. While more
% robust strategies may be conceived, empirically we found that the simpler
% quadratic scheme tends to work reasonably well.
%\paolo{Maybe a plot of an experiment with increased noise in user response? (for example, $\beta=0.1$ instead of $1$.}

@@

We presented a max-margin approach for efficient preference
elicitation in large configuration spaces.\footnote{Note that max-margin learning has been proposed before \cite{gajos2005} for preference elicitation, 
but with  rudimental methods for query selection.}
Our approach relies on an
extension of max-margin learning to sets, and is effective in the
generation of a diverse set of configurations that can be used to ask
informative preference queries.  The main advantages of this
elicitation method are 1) ability to provide recommendations in large
configuration problems 2) robustness with respect to erroneous
feedback and 3) ability to encourage sparse utility functions.
Experimental comparisons against state-of-the-art Bayesian preference
elicitation strategies confirm these advantages. Future work includes
extending the approach to truly hybrid scenarios (where real valued
attributes do not depend on categorical ones) and studying its
applicability to other problems, \paolo{(as the identification of Choquet models \cite{Pine2013}}.

%\section*{Acknowledgments}
\paragraph{Acknowledgments}
ST was supported by the Caritro Foundation through project E62I15000530007.
PV was supported by the Idex Sorbonne Universit\'{e}s under grant ANR-11-IDEX-0004-02.
We thank Craig Boutilier for motivating discussion on the topic.
